### 2012-01-23 docs/SDC/sdc.tex
latex fixes
define I's for enthalpy

\begin{figure}[tb]
\centering
\includegraphics[scale=0.65]{\sdcfigpath/flowchart_SDC}
\caption[Graphical flowchart of \maestro\ SDC]
  {\label{fig:sdc:flowchart} A flowchart of the \maestro SDC algorithm.  The
  thermodynamic state variables and local velocity are
  indicated in each step.  The base state is not shown as it is time-independent.
  Red text indicates that quantity was
  updated during that step.  The predictor is 
  outlined by the dotted box.  The blue text indicates state
  variables that are the same in {\bf Step 6} as they are in
  {\bf Step 2}, i.e., they are unchanged by the predictor steps.
  The SDC loop is shown in the gray dotted box.}
\end{figure}

@@

\paragraph{Species Source Terms.}
For the species prediction, the form of $I_{\rho X_k}$ depends on
{\tt species\_pred\_type} (see \S \ref{sec:pred:density}):  
\begin{itemize}
\item {\tt species\_pred\_type} = 1 (predict $\rho'$ and $X_k$)
%
\begin{equation}
I_{\rho X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
      \frac{(\rho X_k)^\mathrm{new} - 
            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
\end{equation}

\item {\tt species\_pred\_type} = 2 (predict $\rho'$ and $(\rho X_k)$)
%
\begin{equation}
I_{\rho X_k} = \frac{(\rho X_k)^\mathrm{new} - (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k} 
\end{equation}

\item {\tt species\_pred\_type} = 3 (predict $\rho$ and $X_k$)
%
\begin{equation}
I_{\rho X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
      \frac{(\rho X_k)^\mathrm{new} - 
            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
\end{equation}

@@

\paragraph{Enthalpy Source Terms.}
For the enthalpy prediction, the form of $I_{\rho h}$ depends on
{\tt enthalpy\_pred\_type} (see \S \ref{sec:pred:enthalpy}).:
\begin{itemize}
\item {\tt enthalpy\_pred\_type} = 1 \\
not right -- should be pert form
if we do something in perturbational form, then don't we need to have
the base state see the reaction term in its prediction?
\begin{equation}
I_{\rho h}  = \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h}
\end{equation}
\end{itemize}

>>>>

\begin{figure}[tb]
\centering
\includegraphics[scale=0.65]{\sdcfigpath/flowchart_SDC}
\caption[Graphical flowchart of \maestro\ SDC]
  {\label{fig:sdc:flowchart} A flowchart of the \maestro\ SDC algorithm.  The
  thermodynamic state variables and local velocity are
  indicated in each step.  The base state is not shown as it is time-independent.
  Red text indicates that quantity was
  updated during that step.  The predictor is 
  outlined by the dotted box.  The blue text indicates state
  variables that are the same in {\bf Step 6} as they are in
  {\bf Step 2}, i.e., they are unchanged by the predictor steps.
  The SDC loop is shown in the gray dotted box.}
\end{figure}

@@

\paragraph{Species Source Terms.}
For the species prediction, the form of $I_{\rho X_k}$ depends on
{\tt species\_pred\_type} (see \S \ref{sec:pred:density}):  
\begin{itemize}
\item {\tt species\_pred\_type} = 1 (predict $\rho'$ and $X_k$)
%
\begin{equation}
I_{X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
      \frac{(\rho X_k)^\mathrm{new} - 
            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
\end{equation}

\item {\tt species\_pred\_type} = 2 (predict $\rho'$ and $(\rho X_k)$)
%
\begin{equation}
I_{\rho X_k} = \frac{(\rho X_k)^\mathrm{new} - (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k} 
\label{eq:sdc:Irhoo}
\end{equation}

\item {\tt species\_pred\_type} = 3 (predict $\rho$ and $X_k$)
%
\begin{equation}
I_{X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
      \frac{(\rho X_k)^\mathrm{new} - 
            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
\end{equation}

@@

\paragraph{Enthalpy Source Terms.}
The obvious analog to the above form for $I_{\rho X_k}$ is to define
\begin{equation}
I_{\rho h}  = \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h}
\label{eq:sdc:Irhoh}
\end{equation}
This can be thought of as an approximation to $(\rho \Hnuc)$,
but this does not match any of the quantities we predict for the
enthalpy update ({\tt enthalpy\_pred\_type}, see \S
\ref{sec:pred:enthalpy}).  The appropriate constructions
are:

\begin{itemize}
\item {\tt enthalpy\_pred\_type} = {\tt predict\_rhoprime}

Not implemented.

Here we need an $I_{\rho h}$ term for the $(\rho h)'$ evolution
equation (see Eq.~\ref{rhohprime equation}).  Defining the source
terms for this requires a base state enthalpy that knows about the
source terms---this is currently not implemented.

In particular, should $(\rho h)_0$ see a $\overline{(\rho \Hnuc)}$ term?
what about an average thermal diffusion?

\item {\tt enthalpy\_pred\_type} = {\tt predict\_h}

This is the most straightforward prediction type.  The SDC solver
integrates the equation for $(\rho h)$:
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt}  + \rho H_{\rm nuc} 
\end{equation}
(shown here without diffusion or external heat sources).  Expanding
the time derivative and divergence, and using the continuity equation
we see:
\begin{equation}
\frac{\partial h}{\partial t} = -\Ub \cdot \nabla h + \frac{1}{\rho} \frac{Dp_0}{Dt}  + \frac{1}{\rho} (\rho H_{\rm nuc}) \label{eq:sdc:h}
\end{equation}
Comparing these equations, we see that 
\begin{equation}
I_{h}  = \frac{1}{\rho^{n+\myhalf}} \left [
    \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h} \right ]
\end{equation}

\item {\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_rhoprime} or 
      {\tt enthalpy\_pred\_type} = {\tt predict\_T\_then\_h}

Both of these {\tt enthalpy\_pred\_type}s predict temperature.  Expressing
$h = h(p_0,T,X_k)$ and differentiating along particle paths:
\begin{eqnarray}
\frac{Dh}{Dt} &=& \left . \frac{\partial h}{\partial T} \right |_{p,X_k} \frac{DT}{Dt} +
                  \left . \frac{\partial h}{\partial p} \right |_{T,X_k} \frac{Dp_0}{Dt} +
           \sum_k \left . \frac{\partial h}{\partial X_k} \right |_{p,T} \frac{DX_k}{Dt} \\
              &=& c_p \frac{DT}{Dt} + h_p  \frac{Dp_0}{Dt} + \sum_k \xi_k \omegadot_k
\end{eqnarray}
where $c_p$, $h_p$, and $\xi_k$ are as defined in the table of symbols
(Table~\ref{table:sym}), and we substitute $DX_k/Dt = \omegadot_k$ (from the species
continuity equation, Eq.~\ref{species equation}).  Using Eq.~\ref{eq:sdc:h}, we have
the familiar temperature evolution equation:
\begin{equation}
\rho c_p \frac{DT}{Dt} = \underbrace{(1 - \rho h_p) \frac{Dp_0}{Dt}}_{\begin{smallmatrix}\text{already~accounted~for} \\ \text{in~T~prediction}\end{smallmatrix}} - \sum_k \xi_k \rho \omegadot_k + \rho \Hnuc
\end{equation}
where the underbraced term is already present in {\tt mktempforce}.  Recognizing that
Eq.~\ref{eq:sdc:Irhoh} is the SDC approximation to $(\rho \Hnuc)$ and Eq.~\ref{eq:sdc:Irhoo} is the 
SDC approximation to $(\rho \omegadot_k)$, we can define
\begin{equation}
I_T = \frac{1}{\rho^{n+\myhalf} c_p^{n+\myhalf}} \left \{
  \left [ \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h} \right ]
  - \sum_k \xi_k^{n+\myhalf} \left [      \frac{(\rho X_k)^\mathrm{new} - 
            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ] \right \}
\end{equation}
